Repeatability of Pre-Flashover Fire Patterns on Gypsum Wallboard

### 2015-05-11 Projects/Madrzykowski_Thesis/Report/Madrzykowski_Thesis.tex
minor edits to intro

\acknowledgments{
I would like to thank Prof.~Fleishmann, Dr.~Kevin McGrattan, and Dr.~Craig Weinschenk from the National Institute of Standards and Technology for their guidance, assistance, and support throughout this study.

@@

Comparative experiments were conducted to examine the impact of the method of construction and support of the walls and corner sections on the development of the burn patterns.  Three cases were examined: (1) 12.7~mm thick, regular core, painted gypsum board on the front of a wall frame (open back); (2) 12.7~mm thick, regular core, painted gypsum board on the front and back of a wall frame (interior wall); and {3) 12.7~mm thick, regular core, painted gypsum board on the front and back of a wall frame, with fiberglass batting insulation in the voids of the wood frame (exterior wall).  The main consideration was to identify any differences in the height, width, shape, and area of the fire patterns. The results of these experiments determined the construction of the walls and corners used in subsequent phases of this study.

@@

A series of experiments were conducted with samples of painted gypsum wallboard to determine the repeatability of a ``critical heat flux'' that results in a clear line of demarcation, a ``burn/no burn'' line.  The flooring radiant panel (ASTM~E648) was used to examine this value.

@@

Cone calorimeter~\cite{ASTM_E1354} experiments were conducted to estimate the ignition temperature, the time to ignition and the heat release rate from samples of painted gypsum wallboard under thermal flux conditions ranging from 25~kW/m$^2$ to 75~kW/m$^2$.

@@

The ability to recreate or simulate a fire pattern or certain characteristics of a fire pattern was examined with a variety of tools.  The first characteristic examined was the ability to determine based on the calculated heat flux to the wall if a ``visible or measurable physical change'' would occur to the painted paper surface of the gypsum wallboard.  The next characteristic of the fire pattern that was examined was the height.  The height of the fire pattern was then compared to the predictive results from flame height correlations. Heskestad developed an empirical method for calculating the mean visible flame height for fuels which do not have substantial in-depth combustion~\cite{Beyler:1986,Heskestad:SFPE}. This correlation was compared with the flame height results of the source fires and the instrumented wall experiments.

>>>>

\acknowledgments{
I would like to thank Prof.~Fleishmann, my senior advisor with the University of Canterbury, as well as, Dr.~Kevin McGrattan, and Dr.~Craig Weinschenk from the National Institute of Standards and Technology for their guidance, assistance, and support throughout this study.

@@

Comparative experiments were conducted to examine the impact of the method of construction and support of the walls and corner sections on the development of the burn patterns.  Three cases were examined: (1) 12.7~mm thick, regular core, painted gypsum wallboard on the front of a wall frame (open back); (2) 12.7~mm thick, regular core, painted gypsum wallboard on the front and back of a wall frame (interior wall); and {3) 12.7~mm thick, regular core, painted gypsum wallboard on the front and back of a wall frame, with fiberglass batting insulation in the voids of the wood frame (exterior wall).  The main consideration was to identify any differences in the height, width, shape, and area of the fire patterns. The results of these experiments determined the construction of the walls and corners used in subsequent phases of this study.

@@

A series of experiments were conducted with samples of painted gypsum wallboard to determine the repeatability of a ``critical heat flux'' that results in a clear line of demarcation, a ``burn/no burn'' line.  The flooring radiant panel, ASTM~E648, and a modified radiant panel apparatus based on ASTM~E162 were used to examine this value~\cite{ASTM_E648},~\cite{ASTM_E162}.

@@

Cone calorimeter, ASTM~E1354, experiments were conducted to estimate the ignition temperature, the time to ignition and the heat release rate from samples of painted gypsum wallboard under thermal flux conditions ranging from 25~kW/m$^2$ to 75~kW/m$^2$ ~\cite{ASTM_E1354}.

@@

The ability to recreate or simulate a fire pattern or certain characteristics of a fire pattern was examined with a variety of tools.  The first characteristic examined was the ability to determine, based on the calculated heat flux to the wall, if a ``visible or measurable physical change'' would occur to the painted paper surface of the gypsum wallboard.  The next characteristic of the fire pattern that was examined was the height.  The height of the fire pattern was then compared to the predictive results from flame height correlations. Heskestad developed an empirical method for calculating the mean visible flame height for fuels which do not have substantial in-depth combustion~\cite{Beyler:1986,Heskestad:SFPE}. This correlation was compared with the flame height results of the source fires and the instrumented wall experiments.
